\documentclass[runningheads]{llncs}

% Packages
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,fit,backgrounds}

% Custom commands
\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\bigO}{\mathcal{O}}

% Code listing settings
\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=flexible,
  keepspaces=true
}

\begin{document}
\titlerunning{EVMAuth}
\title{EVMAuth: Decentralized Authorization State Management for Web3 APIs and AI Agents}

\author{Shaun Scovil\inst{1} \and Bhargav Chickmagalur Nanjundappa\inst{1}}

\institute{
Radius Technology Systems, Cambridge, USA\\
\email{\{shaun,bhargav\}@radius.tech}
}

\maketitle

\begin{abstract}
Traditional API authorization requires complex infrastructure: OAuth servers, databases for session management, subscription services, and payment processors.
This creates operational overhead, vendor lock-in, and single points of failure.
We present EVMAuth, a complete on-chain authorization framework that eliminates these dependencies by using blockchain token ownership as the authorization primitive.
EVMAuth provides seven composable features: role-based access control, account freezing, token enumeration, time-to-live expiration, transfer control, direct purchase, and dynamic configuration.
We implement EVMAuth in both ERC-1155 and ERC-6909 token standards and deploy to Base testnet.
Our evaluation shows that EVMAuth reduces authorization infrastructure from multiple services to a single smart contract while maintaining sub-100ms verification latency.
Gas costs scale with O(k) for most operations and O(k²) for worst-case transfers, where k is a configurable storage bound.
We demonstrate real-world applicability through an API gateway integration and discuss implications for AI agent authentication.

\keywords{Blockchain \and Authorization \and Smart contracts \and API security \and Decentralized systems \and AI agents}
\end{abstract}

% Include sections
\input{sections/01-introduction}
\input{sections/02-background}
\input{sections/03-design}
\input{sections/04-implementation}
\input{sections/05-evaluation}
\input{sections/06-discussion}
\input{sections/07-related}
\input{sections/08-conclusion}

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
